Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-20 ba_resubmit/kundinger_final.tex
done with most edits from jerry.
update.
cleaning up comments.

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  \textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
%
%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}

It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
\begin{align}
\label{eqn:sadinle_prior}
p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.

@@

\begin{enumerate}	
\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined. Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$. It seems that the conditioning may actually be important here. I mention this for two reasons, it appears a bit strange in equation (22). When I think through it, it's okay. I just wanted to bring it up as a conversation before submission.}
\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1. Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly. Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
\end{enumerate}

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$. We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.

>>>>

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  \textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
%
%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
%
It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
\begin{align}
\label{eqn:sadinle_prior}
p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.

@@

%\begin{enumerate}	
%\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined. Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$. It seems that the conditioning may actually be important here. I mention this for two reasons, it appears a bit strange in equation (22). When I think through it, it's okay. I just wanted to bring it up as a conversation before submission.}
%\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
%\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1. Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly. Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
%\end{enumerate}

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$. We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.
